% BIEGrid benchmark summary
\documentclass[11pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}

\newcommand{\improved}[1]{%
  \begingroup
  \setlength{\fboxsep}{1pt}%
  \colorbox{green!20}{\strut #1}%
  \endgroup
}

\begin{document}

\section*{BIEGrid Benchmark Summary ($k=p=40$)}

Square BIEGrid, double precision, $\kappa=0$, one right-hand side,
GMRES tolerance $10^{-10}$, \texttt{num\_threads=4}, and
\texttt{save\_samples=false}, \texttt{load\_samples=false}.
Reported quantities:
\[
\text{relerr}=\frac{\|A-A_{\mathrm{rsrs}}\|_2}{\|A\|_2},
\qquad
\text{errsolve}=\|I-A_{\mathrm{rsrs}}^{-1}A\|_2.
\]
The $\mathrm{cond}(A)$ column uses the PerLevel estimate.

\begin{table}[h]
\centering
\small
\begin{tabular}{r r r r r r r r}
\hline
& & \multicolumn{3}{c}{PerLevel ($k=p=40$)} & \multicolumn{3}{c}{Constant ($k=p=40$)} \\
\cline{3-5} \cline{6-8}
$N$ & $\mathrm{cond}(A)$ & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve \\
\hline
80,656  & 1.4e+05 & 1,640 & 2.8e-06 & 1.0e-01 & 1,640 & 3.1e-06 & 1.0e-01 \\
160,801 & 2.8e+05 & 1,640 & 8.8e-06 & 5.5e-01 & 1,640 & 8.4e-06 & 5.4e-01 \\
321,489 & 7.5e+05 & 1,640 & 2.8e-05 & 4.2e+00 & --    & --      & --      \\
\hline
\end{tabular}
\caption{Requested stop-point comparison for completed $k=p=40$ runs for \texttt{bempp/rsrs} commit \texttt{efe95a4}.}
\label{tab:biegrid_k40_stop_point}
\end{table}

\section*{Fixed Rank Sampling Rules}

\noindent\fbox{%
\parbox{0.97\textwidth}{%
\textbf{Sampling rule.}\\[2pt]
1. \textbf{Upward pass through the tree:} if $B$ is a leaf box, set
\[
|B|_{\mathrm{act}}=\#\{\text{points in }B\}.
\]
If $P$ is a coarser box with children $\mathcal C(P)$, set
\[
|P|_{\mathrm{act}}
=
\sum_{C\in\mathcal C(P)} \min\!\bigl(k,|C|_{\mathrm{act}}\bigr).
\]
For each non-root box $B$ on level $\ell$, define
\[
|N(B)|_{\mathrm{act}}
=
\sum_{B'\in N_\ell(B)} |B'|_{\mathrm{act}},
\]
and then compute
\[
s_B
=
|B|_{\mathrm{act}}
+
|N(B)|_{\mathrm{act}}
+
\min\!\bigl(k,|B|_{\mathrm{act}}\bigr)
+
p,
\]
and store this value using the pair $(\ell,\mathrm{Morton}(B))$.

\medskip\noindent
2. After the upward pass is complete, accumulate the remaining skeleton
indices at the root and set
\[
s_{\max}=\max_{B \neq \mathrm{root}} s_B,
\qquad
s_0=\min\!\bigl(s_{\max},\,2\,\# I_{S,0}+2p\bigr),
\]
where $I_{S,0}$ denotes the accumulated skeleton block at the end of the
upward sweep.

\medskip\noindent
3. \textbf{Constant and PerLevel:} in both modes, the maintained sketch data
have width $s_{\max}$ throughout the upward sweep. The terminal root block
uses the cap $s_0$ from point 2.

\medskip\noindent
4. In \textbf{Constant}, the box-local nullspace and pseudoinverse
computations use $s_{\max}$. In \textbf{PerLevel}, the code retrieves the
saved value $s_B$ by the current level and the current box Morton key and uses
that boxwise count in these local computations. Thus the distinction between
the two modes lies in the number of samples used for the linear-algebraically
intensive local computations, while the maintained sketch data have width
$s_{\max}$ in both cases.%
}}

\paragraph{Remark.}
Once the factorization begins, the stored sketch data of width $s_{\max}$ are
updated under every ID and LU action on every level. In other words, the
columns corresponding to the stored samples are propagated through the
accumulated $E$, $F$, $L$, and $U$ transformations as they are formed. Hence
the maintained sketch data are already current at the end of the upward sweep,
and no second traversal is required later to replay these transformations onto
the stored samples. In the fixed-rank setting, the box-local nullspace and
pseudoinverse calculations still use the saved count $s_B$ associated with the
current box.

\paragraph{Remark.}
During the RSRS factorization, indices are progressively marked
inactive as each level advances through its eliminations. The upward-pass rule
in point 1 does not recompute $s_B$ from these evolving inactive sets within a
level; instead it keeps the precomputed boxwise count. This is deliberate,
since using the dynamically reduced active neighbor counts in the local
pseudoinverse calculations increased the error.

\paragraph{Benchmark Experiments.}
For the runs below we use square BIEGrid in double precision with
$\kappa=0$, one right-hand side, GMRES tolerance $10^{-10}$,
\texttt{num\_threads=4}, deterministic seed 12345, and no saved sampling cache.
Table~2 reports the current \texttt{bempp/rsrs} baseline after the amended
\texttt{enable\_symm} commit. For the $N=321{,}489$ row we omit the auxiliary
right-hand-side solve and report only the quantities that appear in the table.
Here $k$ denotes the fixed-rank parameter and $p$ the oversampling parameter.

\begin{table}[h]
\centering
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{r r r r r r r r r r r r r}
\hline
& \multicolumn{3}{c}{PerLevel ($k=p=40$)} & \multicolumn{3}{c}{Constant ($k=p=40$)} & \multicolumn{3}{c}{PerLevel ($k=p=60$)} & \multicolumn{3}{c}{Constant ($k=p=60$)} \\
\cline{2-4} \cline{5-7} \cline{8-10} \cline{11-13}
$N$ & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve \\
\hline
80,656  & 1,520 & \improved{6.55e-07} & \improved{2.17e-02} & 1,520 & \improved{4.79e-07} & \improved{1.61e-02} & 2,280 & 4.23e-08 & 1.92e-03 & 2,280 & 3.90e-08 & 1.70e-03 \\
160,801 & 1,520 & \improved{9.10e-07} & \improved{5.00e-02} & 1,520 & \improved{8.21e-07} & \improved{4.81e-02} & 2,280 & 8.38e-08 & 6.78e-03 & 2,280 & 1.28e-07 & 9.15e-03 \\
321,489 & 1,520 & \improved{1.68e-06} & \improved{1.42e-01} & 1,520 & 1.30e-06 & 1.32e-01 & --    & --      & --      & --    & --      & --      \\
\hline
\end{tabular}
}
\caption{Refreshed BIEGrid comparison for \texttt{bempp/rsrs} commit \texttt{1ca2c90}. Green entries indicate substantial improvement relative to the corresponding values in Table~1. Blank entries mark cases that have not yet been rerun for this pinned commit.}
\label{tab:biegrid_perlevel_constant_80k_160k}
\end{table}

\section*{Parallel processing by color}

At each level $\ell$, the implementation first forms the set of boxes that are
eligible to launch. In the fixed-rank setting this means
\[
\mathcal B_\ell
=
\left\{
B \;:\; |I_B|_{\mathrm{act}} > k
\right\},
\]
while in the general case only empty boxes are excluded. The coloring step is
then applied only to the launchable set $\mathcal B_\ell$.

\noindent\fbox{%
\parbox{0.97\textwidth}{%
\textbf{Coloring rule.}\\[2pt]
For each level $\ell$, initialize all launchable boxes as uncolored and sweep
through the boxes in the current upward traversal order:
\begin{enumerate}
\item for each launchable box $B \in \mathcal B_\ell$:
\item collect the geometric neighbor list $\widetilde N_\ell(B)$ returned by the tree search;
\item among these neighbors, keep only those boxes that have already been encountered in the sweep and already carry a color;
\item note that $\widetilde N_\ell(B)$ may also contain coarse boxes above $B$ that have not yet been assigned a color in the present sweep, and these contribute no color constraint at this stage;
\item assign to $B$ the smallest nonnegative color that is not already used by a previously colored box whose near-field list overlaps that of $B$.
\end{enumerate}
Equivalently, if $N_\ell^{\mathrm{past}}(B)$ denotes the subset of neighbors of
$B$ that have already been encountered and assigned a color, and if
\[
\mathcal O_\ell(B)
=
\left\{
B' \in N_\ell^{\mathrm{past}}(B)
\;:\;
\widetilde N_\ell(B') \cap \widetilde N_\ell(B) \neq \emptyset
\right\},
\]
then
\[
\mathrm{color}(B)
=
\min\Bigl\{
c \in \{0,1,2,\dots\}
\;:\;
c \notin
\{
\mathrm{color}(B') : B' \in \mathcal O_\ell(B)
\}
\Bigr\}.
\]
Thus any two boxes that share a near-field neighbor are assigned different
colors. Consequently, any two boxes of the same color have nonoverlapping near
fields and can be processed in parallel on that level.%
}}

For a fixed color class $\mathcal C_{\ell,c}$, all boxes in
$\mathcal C_{\ell,c}$ are processed in parallel. The code first performs the
box-local nullspace / ID step for every $B \in \mathcal C_{\ell,c}$ in
parallel, updates the maintained sketch data under the resulting ID factors,
and then performs the box-local LU / pseudoinverse step for the boxes that
remain active after the ID stage. Once the LU stage for the color is complete,
the newly eliminated residual indices are appended to the inactive set and the
near-field index lists are filtered before the next color is processed.

\paragraph{Remark.}
The launch mask removes boxes for which the fixed-rank local computation would
be vacuous, namely boxes with $|I_B|_{\mathrm{act}} \le k$. Such boxes are not
colored and do not enter the per-color parallel work queue. This keeps only
meaningful local computations that actually launch, and avoids carrying
degenerate nullspace or pseudoinverse calls through the parallel pipeline.

\begin{table}[h]
\centering
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{r r r r r r r r r r r r r}
\hline
& \multicolumn{3}{c}{PerLevel ($k=p=40$)} & \multicolumn{3}{c}{Constant ($k=p=40$)} & \multicolumn{3}{c}{PerLevel ($k=p=60$)} & \multicolumn{3}{c}{Constant ($k=p=60$)} \\
\cline{2-4} \cline{5-7} \cline{8-10} \cline{11-13}
$N$ & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve \\
\hline
80,656  & 1,520 & \improved{3.09e-07} & \improved{1.05e-02} & 1,520 & \improved{2.67e-07} & \improved{9.51e-03} & 2,280 & 3.50e-08 & 1.55e-03 & 2,280 & \improved{2.35e-08} & \improved{8.56e-04} \\
160,801 & 1,520 & 6.54e-07 & 4.07e-02 & 1,520 & \improved{5.61e-07} & \improved{3.00e-02} & 2,280 & 1.27e-07 & 8.58e-03 & 2,280 & 1.06e-07 & 7.10e-03 \\
321,489 & 1,520 & \improved{1.15e-06} & \improved{9.58e-02} & 1,520 & \improved{5.68e-07} & \improved{5.12e-02} & --    & --      & --      & --    & --      & --      \\
\hline
\end{tabular}
}
\caption{Refreshed BIEGrid comparison for \texttt{bempp/rsrs} commit \texttt{21470dd}. Green entries indicate substantial improvement relative to the corresponding values in Table~2. Blank entries mark cases that have not yet been rerun for this pinned commit.}
\label{tab:biegrid_perlevel_constant_21470dd}
\end{table}

\pagebreak{}
\section*{Diagonal Extraction in Joint ID and LU stage}

For commit \texttt{ee38df9}, the box-local diagonal block associated with a
launched LU computation is no longer treated as data that must be extracted a
second time later from the maintained sketches. Instead, when the joint ID/LU
step forms the local $RR$ block for a box $B$, that same block is used
immediately in the diagonal factor construction. In the real symmetric route
this is the local $X_{RR}$ block itself, while in the nonsymmetric route the
code forms a symmetrized $X_{RR}$ by extracting the corresponding local
$RR$ views from the maintained $Y$ and $Z$ sketch data and averaging them
before factorization.

\paragraph{Remark.}
The stored diagonal data are also made leaner in this commit. Rather than
retaining both the dense block $X_{RR}$ and a second factorization
$\mathrm{LU}(X_{RR})$, the implementation stores only the factorized LU
representation needed later by the RSRS apply and solve routines. In an
earlier version, maintaining both $X_{RR}$ and $\mathrm{LU}(X_{RR})$ was
helpful for maintaining the adjoint check. With the present implementation of
diagonal extraction from the joint ID/LU stage, this redundancy is no longer
necessary. The final accumulated skeleton block at the root is still treated
as a terminal computation performed at the very end of the RSRS factorization.

\begin{table}[h]
\centering
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{r r r r r r r r r r r r r}
\hline
& \multicolumn{3}{c}{PerLevel ($k=p=40$)} & \multicolumn{3}{c}{Constant ($k=p=40$)} & \multicolumn{3}{c}{PerLevel ($k=p=60$)} & \multicolumn{3}{c}{Constant ($k=p=60$)} \\
\cline{2-4} \cline{5-7} \cline{8-10} \cline{11-13}
$N$ & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve & $n_{\mathrm{samples}}$ & relerr & errsolve \\
\hline
80,656  & 1,520 & 3.09e-07 & 1.12e-02 & 1,520 & 2.51e-07 & 1.01e-02 & 2,280 & 4.58e-08 & 1.91e-03 & 2,280 & 2.74e-08 & 1.13e-03 \\
160,801 & 1,520 & 7.29e-07 & 3.57e-02 & 1,520 & 6.07e-07 & 3.51e-02 & 2,280 & 9.37e-08 & 6.95e-03 & 2,280 & 9.02e-08 & 6.64e-03 \\
321,489 & 1,520 & 9.69e-07 & 8.89e-02 & 1,520 & 5.81e-07 & 5.02e-02 & -- & -- & -- & -- & -- & -- \\
\hline
\end{tabular}
}
\caption{BIEGrid comparison for \texttt{bempp/rsrs} commit \texttt{ee38df9}. The $N=321{,}489$, $k=p=60$ entries are intentionally omitted.}
\label{tab:biegrid_perlevel_constant_ee38df9}
\end{table}


\end{document}
