\documentclass[11pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath}
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}

\title{Perturbed BIEGrid RSRS Regression Changelog}
\author{}
\date{\today}

\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
\newcommand{\fail}[1]{\cellcolor{orange!30}{#1}}
\newcommand{\ok}[1]{#1}

\begin{document}
\maketitle

\section*{Setup}

All runs below are serial 10k BIEGrid RSRS checks: mesh width \texttt{0.01},
10201 points, fixed rank \texttt{k = 40}, oversampling \texttt{p = 40}, and
\texttt{tot\_num\_samples = 1520}. The RSRS sampling seed, perturbation seed,
and error diagnostic seeds are fixed. Seeding initializes the random number
generator at the start of a run; it does not reuse the same sample vector
repeatedly. Each draw advances the generator, and repeating the run with the
same seed reproduces the same sample sequence.

The adjoint check is
\[
    \ip{Ax}{y} \approx \ip{x}{A^*y}.
\]
It is applied to the RSRS approximation and to the inverse/preconditioner path.
When this check fails, the relative 2-norm estimate can become larger than the
relative Frobenius norm estimate.

\section*{Perturbed Operator}

Let $A_0$ denote the unperturbed real symmetric BIEGrid operator. The perturbed
operators have the form $A = A_0 + \Delta$. The vectors $u$ and $v$ are
deterministic random vectors generated from perturbation seed \texttt{12345} and
normalized to have 2-norm one. The perturbation scale is denoted by $s$. For
real perturbations, $\alpha = s$. For complex perturbations,
$\alpha = s(1+2i)/\sqrt{5}$, so $|\alpha| = s$.

For nonsymmetric perturbations,
\[
    \Delta = \alpha u v^*,\qquad
    \Delta x = \alpha u (v^* x),\qquad
    \Delta^* x = \overline{\alpha}\, v (u^* x).
\]
For real data, $v^* = v^T$ and $\overline{\alpha} = \alpha$. For the real
symmetric variant, $v = u$. For the complex symmetric variant,
\[
    \Delta = \alpha u \overline{u}^*,\qquad
    \Delta x = \alpha u (\overline{u}^* x),\qquad
    \Delta^* x = \overline{\alpha}\,\overline{u}(u^* x).
\]
The exact perturbed operator tests compare these \texttt{matvec} and
\texttt{rmatvec} routes against the explicitly assembled dense matrix and its
transpose/adjoint. These exact operator adjoint checks pass at roundoff level;
the failures below are for the RSRS approximation and inverse.

\section*{Change}

The old results use \texttt{bempp/rsrs} commit \texttt{4a2fcb2} with
$s=3{\times}10^{-3}$. The new results use commit \texttt{12a0c28} with
$s=10^{-2}$. The algorithmic changes described below are changes to
\texttt{bempp/rsrs}.

The change separates the treatment of local LU factors used for near-field
updates from the diagonal LU factors used in the final block diagonal solve.
Let the local pivot block be $X_{RR}$, and let its partial-pivoted LU factorization be
$P X_{RR} = L U$. Partial pivoting is row pivoting, so the algebra for applying
$X_{RR}^{-1}$ on the left is not equivalent to applying the same triangular
factors from the right or in an adjoint path. For near-field LU updates, the
pivoted solve is now evaluated during extraction to form dense rectangular
blocks such as
\[
    G_{RN} = X_{RR}^{-1} X_{RN}, \qquad
    G_{NR} = X_{NR} X_{RR}^{-1},
\]
and the resulting update blocks are stored directly. For diagonal LU factors,
the unfactorized pivot block $X_{RR}$ is stored alongside the LU factorization,
so non-inverse transpose/adjoint application can use $X_{RR}^*$ directly rather
than applying row-pivoted triangular factors in an incompatible orientation.

This does not add a second rectangular near-field block: the precomputed update
block replaces the previously stored rectangular block, and the square LU factor
for that near-field update is no longer stored. The storage increase is in the
diagonal LU factors, which retain both the unfactorized pivot block $X_{RR}$ and
the LU factorization.

\section*{Old Results: \texttt{4a2fcb2}, $s=3{\times}10^{-3}$}

\begin{center}
\scriptsize
\resizebox{\textwidth}{!}{%
\begin{tabular}{>{\raggedright\arraybackslash}p{1.55in}
                >{\raggedright\arraybackslash}p{0.7in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}}
\hline
Case & RSRS symmetry & $\mathrm{cond}_2(A)$ & rel. 2-norm & rel. Fro. & adjoint check & inverse adjoint & solve error \\
\hline
Unperturbed BIEGrid & \texttt{Symmetric} &
$1.5176{\times}10^{4}$ & $9.3268{\times}10^{-8}$ & $1.7279{\times}10^{-7}$ &
$1.6369{\times}10^{-14}$ & $4.4149{\times}10^{-14}$ & $1.2546{\times}10^{-5}$ \\
Real symmetric perturbed & \texttt{Symmetric} &
$1.5176{\times}10^{4}$ & \fail{$7.2546{\times}10^{-6}$} & $3.3677{\times}10^{-7}$ &
\fail{$1.7151$} & \fail{$1.2221$} & $2.1916{\times}10^{-5}$ \\
Real nonsymmetric perturbed & \texttt{NoSymm} &
$2.4671{\times}10^{6}$ & $2.2110{\times}10^{-7}$ & $6.6145{\times}10^{-7}$ &
$7.8270{\times}10^{-15}$ & $2.3497{\times}10^{-13}$ & $9.8832{\times}10^{-5}$ \\
Complex symmetric perturbed & \texttt{Symmetric} &
$5.2077{\times}10^{5}$ & \fail{$1.1939{\times}10^{-5}$} & $6.8507{\times}10^{-7}$ &
\fail{$4.3644{\times}10^{-1}$} & \fail{$1.2251$} & $1.3907{\times}10^{-4}$ \\
Complex nonsymmetric perturbed & \texttt{NoSymm} &
$9.2613{\times}10^{5}$ & \fail{$8.0830{\times}10^{20}$} & \fail{$1.3552{\times}10^{14}$} &
\fail{$1.0000$} & \fail{$1.0000$} & \fail{$2.2829{\times}10^{17}$} \\
\hline
\end{tabular}
}
\end{center}

\section*{New Results: \texttt{12a0c28}, $s=10^{-2}$}

\begin{center}
\scriptsize
\resizebox{\textwidth}{!}{%
\begin{tabular}{>{\raggedright\arraybackslash}p{1.55in}
                >{\raggedright\arraybackslash}p{0.7in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}
                >{\raggedleft\arraybackslash}p{0.75in}}
\hline
Case & RSRS symmetry & $\mathrm{cond}_2(A)$ & rel. 2-norm & rel. Fro. & adjoint check & inverse adjoint & solve error \\
\hline
Unperturbed BIEGrid & \texttt{Symmetric} &
$1.5176{\times}10^{4}$ & $7.0224{\times}10^{-8}$ & $1.7457{\times}10^{-7}$ &
$4.5513{\times}10^{-14}$ & $3.1644{\times}10^{-14}$ & $1.4197{\times}10^{-5}$ \\
Real symmetric perturbed & \texttt{Symmetric} &
$1.5176{\times}10^{4}$ & $1.3203{\times}10^{-7}$ & $4.4119{\times}10^{-7}$ &
$2.2423{\times}10^{-13}$ & $5.5531{\times}10^{-14}$ & $2.8875{\times}10^{-5}$ \\
Real nonsymmetric perturbed & \texttt{NoSymm} &
$5.1845{\times}10^{7}$ & $1.8230{\times}10^{-7}$ & $6.6715{\times}10^{-7}$ &
$7.6885{\times}10^{-14}$ & $6.5396{\times}10^{-13}$ & $1.7199{\times}10^{-3}$ \\
Complex symmetric perturbed & \texttt{Symmetric} &
$6.1650{\times}10^{5}$ & \fail{$1.8288{\times}10^{-5}$} & $6.5870{\times}10^{-7}$ &
$1.6388{\times}10^{-14}$ & $6.9111{\times}10^{-14}$ & $9.8468{\times}10^{-5}$ \\
Complex nonsymmetric perturbed & \texttt{NoSymm} &
$1.1036{\times}10^{6}$ & \fail{$2.1452{\times}10^{16}$} & \fail{$1.4810{\times}10^{16}$} &
$5.6176{\times}10^{-14}$ & \fail{$1.8260{\times}10^{-1}$} & \fail{$1.7886{\times}10^{18}$} \\
\hline
\end{tabular}
}
\end{center}

\noindent Orange cells indicate the metrics that fail the regression thresholds,
show the remaining large-error behavior, or violate the consistency check
$\text{rel. 2-norm} \le \text{rel. Fro.}$.

\section*{Complex Nonsymmetric Split Check}

The complex nonsymmetric perturbed case still fails as a combined complex
problem. As a diagnostic, the corresponding real and imaginary component
problems were run separately as real problems at $s=10^{-2}$. Both pass the
apply and adjoint checks:
\[
\begin{array}{lrrrr}
\hline
\text{case} & \text{rel. 2-norm} & \text{rel. Fro.} & \text{adjoint check} & \text{solve error}\\
\hline
\text{real part} & 5.2073{\times}10^{-7} & 2.1513{\times}10^{-6} & 7.4251{\times}10^{-14} & 5.6595{\times}10^{-4}\\
\text{imaginary part} & 5.5817{\times}10^{-7} & 2.2066{\times}10^{-6} & 3.7260{\times}10^{-14} & 5.8883{\times}10^{-3}\\
\hline
\end{array}
\]

\end{document}
